\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-08-21 analysis/paper/paper.tex

\caption{\label{tab:print-erspc-estimates}Comparison of estimated hazard ratios and 95\% confidence intervals for ERSPC data.}
\centering
\begin{tabular}[t]{>{\bfseries}lcc}
\toprule
Model & casebase::fitSmoothHazard & survival::survreg\\
\midrule
Exponential & 0.81 (0.68, 0.96) & 0.81 (0.68, 0.96)\\
\cmidrule{1-3}
Gompertz & 0.79 (0.66, 0.94) & 0.80 (0.67, 0.95)\\
\cmidrule{1-3}
Weibull & 0.81 (0.68, 0.96) & 0.80 (0.65, 0.96)\\
\cmidrule{1-3}
Splines & 0.81 (0.68, 0.96) & --\\
\bottomrule
\multicolumn{3}{l}{Cox model estimate: HR (95\% CI) = 0.80 (0.67, 0.95)}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:bmtcrr-cis}Estimates and confidence intervals for the hazard ratios for each coefficient. Both estimates from case-base sampling and Cox regression are presented.}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Case-Base} & \multicolumn{2}{c}{Cox} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Covariates & HR & 95\% CI & HR & 95\% CI\\
\midrule
Sex & 0.72 & (0.42, 1.26) & 0.68 & (0.39, 1.19)\\
Disease & 0.54 & (0.3, 0.97) & 0.52 & (0.29, 0.93)\\
Phase (CR2 vs. CR1) & 1.21 & (0.48, 3.02) & 1.21 & (0.48, 3.02)\\
Phase (CR3 vs. CR1) & 1.63 & (0.41, 6.44) & 1.67 & (0.43, 6.5)\\
Phase (Relapse vs. CR1) & 4.43 & (2.06, 9.55) & 4.55 & (2.09, 9.9)\\
\addlinespace
Source & 1.51 & (0.49, 4.69) & 1.46 & (0.47, 4.54)\\
Age & 0.99 & (0.97, 1.01) & 0.99 & (0.97, 1.02)\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{CodeOutput}
#> Model1 Model2 Model3 
#>    827    793    792
\end{CodeOutput}
\end{CodeChunk}

@@

This report was generated on 2020-08-20 19:21:21 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.2 (2019-12-12)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Pop!_OS 19.10
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/openblas/libblas.so.3
#> LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.7.so
#> 
#> locale:
#>  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
#>  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
#>  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] scales_1.1.1              forcats_0.5.0            
#>  [3] stringr_1.4.0             dplyr_1.0.0              
#>  [5] purrr_0.3.4               readr_1.3.1              
#>  [7] tidyr_1.1.0               ggplot2_3.3.2.9000       
#>  [9] tidyverse_1.3.0           lubridate_1.7.4          
#> [11] visreg_2.7.0              tibble_3.0.1             
#> [13] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [15] flexsurv_1.1.1            survival_3.1-8           
#> [17] pracma_2.2.9              magrittr_1.5             
#> [19] kableExtra_1.1.0          glmnet_4.0-2             
#> [21] Matrix_1.2-18             data.table_1.12.8        
#> [23] cowplot_1.0.0             colorspace_1.4-1         
#> [25] casebase_0.9.0           
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] htmlTable_1.13.3    base64enc_0.1-3     fs_1.3.2           
#>  [7] rstudioapi_0.11     MatrixModels_0.4-1  fansi_0.4.1        
#> [10] mvtnorm_1.0-12      xml2_1.3.0          codetools_0.2-16   
#> [13] knitr_1.29          Formula_1.2-3       jsonlite_1.7.0     
#> [16] broom_0.7.0         cluster_2.1.0       dbplyr_1.4.2       
#> [19] png_0.1-7           compiler_3.6.2      httr_1.4.1         
#> [22] backports_1.1.8     assertthat_0.2.1    cli_2.0.2          
#> [25] acepack_1.4.1       htmltools_0.5.0     quantreg_5.54      
#> [28] tools_3.6.2         gtable_0.3.0        glue_1.4.1         
#> [31] Rcpp_1.0.4.6        cellranger_1.1.0    vctrs_0.3.1        
#> [34] nlme_3.1-143        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.5         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.5       zoo_1.8-7          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.13        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.6          shape_1.4.4        
#> [58] rlang_0.4.6         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-38     htmlwidgets_1.5.1   cmprsk_2.2-9       
#> [64] tidyselect_1.1.0    deSolve_1.27.1      R6_2.4.1           
#> [67] generics_0.0.2      Hmisc_4.3-1         multcomp_1.4-12    
#> [70] DBI_1.1.0           withr_2.2.0         pillar_1.4.4       
#> [73] haven_2.3.1         foreign_0.8-74      mgcv_1.8-31        
#> [76] nnet_7.3-12         mstate_0.2.12       modelr_0.1.5       
#> [79] crayon_1.3.4        rmarkdown_2.3       timereg_1.9.4      
#> [82] jpeg_0.1-8.1        grid_3.6.2          readxl_1.3.1       
#> [85] reprex_0.3.0        digest_0.6.25       webshot_0.5.2      
#> [88] numDeriv_2016.8-1.1 stats4_3.6.2        munsell_0.5.0      
#> [91] viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    fix-cs1 /home/sahir/git_repositories/cbpaper
#> Remote:   fix-cs1 @ origin (https://github.com/sahirbhatnagar/cbpaper.git)
#> Head:     [f618e21] 2020-08-18: time dependent hazard ratio for ERSPC
\end{CodeOutput}
\end{CodeChunk}

>>>>

\caption{\label{tab:print-erspc-estimates}Comparison of estimated hazard ratios and 95\% confidence intervals for ERSPC data.}
\centering
\begin{tabular}[t]{>{\bfseries}lcc}
\toprule
Model & casebase::fitSmoothHazard & survival::survreg\\
\midrule
Exponential & 0.79 (0.67, 0.94) & 0.81 (0.68, 0.96)\\
\cmidrule{1-3}
Gompertz & 0.80 (0.67, 0.95) & 0.80 (0.67, 0.95)\\
\cmidrule{1-3}
Weibull & 0.81 (0.68, 0.96) & 0.80 (0.65, 0.96)\\
\cmidrule{1-3}
Splines & 0.79 (0.66, 0.94) & --\\
\bottomrule
\multicolumn{3}{l}{Cox model estimate: HR (95\% CI) = 0.80 (0.67, 0.95)}\\
\end{tabular}
\end{table}

@@

\caption{\label{tab:bmtcrr-cis}Estimates and confidence intervals for the hazard ratios for each coefficient. Both estimates from case-base sampling and Cox regression are presented.}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Case-Base} & \multicolumn{2}{c}{Cox} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Covariates & HR & 95\% CI & HR & 95\% CI\\
\midrule
Sex & 0.74 & (0.43, 1.3) & 0.68 & (0.39, 1.19)\\
Disease & 0.52 & (0.29, 0.94) & 0.52 & (0.29, 0.93)\\
Phase (CR2 vs. CR1) & 1.20 & (0.48, 3) & 1.21 & (0.48, 3.02)\\
Phase (CR3 vs. CR1) & 1.59 & (0.41, 6.17) & 1.67 & (0.43, 6.5)\\
Phase (Relapse vs. CR1) & 4.09 & (1.9, 8.81) & 4.55 & (2.09, 9.9)\\
\addlinespace
Source & 1.70 & (0.55, 5.23) & 1.46 & (0.47, 4.54)\\
Age & 0.99 & (0.97, 1.02) & 0.99 & (0.97, 1.02)\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{CodeOutput}
#> Model1 Model2 Model3 
#>    827    791    790
\end{CodeOutput}
\end{CodeChunk}

@@

This report was generated on 2020-08-21 16:53:49 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.2 (2019-12-12)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Pop!_OS 19.10
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/openblas/libblas.so.3
#> LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.7.so
#> 
#> locale:
#>  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
#>  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
#>  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.7.0               dotwhisker_0.5.0         
#>  [3] timereg_1.9.4             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.0              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.1.0               ggplot2_3.3.2.9000       
#> [13] tidyverse_1.3.0           lubridate_1.7.4          
#> [15] visreg_2.7.0              tibble_3.0.1             
#> [17] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.1-8           
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_4.0-2             
#> [25] Matrix_1.2-18             data.table_1.12.8        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0           
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] ggstance_0.3.3      htmlTable_1.13.3    base64enc_0.1-3    
#>  [7] fs_1.3.2            rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.0-12     
#> [13] xml2_1.3.0          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.7.0      cluster_2.1.0      
#> [19] dbplyr_1.4.2        png_0.1-7           compiler_3.6.2     
#> [22] httr_1.4.1          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.54       tools_3.6.2         gtable_0.3.0       
#> [31] Rcpp_1.0.4.6        cellranger_1.1.0    vctrs_0.3.1        
#> [34] nlme_3.1-143        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.5         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.5       zoo_1.8-7          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.13        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.6          shape_1.4.4        
#> [58] rlang_0.4.6         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-38     labeling_0.3        htmlwidgets_1.5.1  
#> [64] cmprsk_2.2-9        tidyselect_1.1.0    deSolve_1.27.1     
#> [67] plyr_1.8.6          R6_2.4.1            generics_0.0.2     
#> [70] Hmisc_4.3-1         multcomp_1.4-12     DBI_1.1.0          
#> [73] withr_2.2.0         pillar_1.4.4        haven_2.3.1        
#> [76] foreign_0.8-74      mgcv_1.8-31         nnet_7.3-12        
#> [79] mstate_0.2.12       modelr_0.1.5        crayon_1.3.4       
#> [82] rmarkdown_2.3       jpeg_0.1-8.1        grid_3.6.2         
#> [85] readxl_1.3.1        reprex_0.3.0        digest_0.6.25      
#> [88] webshot_0.5.2       numDeriv_2016.8-1.1 stats4_3.6.2       
#> [91] munsell_0.5.0       viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    fix-cs1 /home/sahir/git_repositories/cbpaper
#> Remote:   fix-cs1 @ origin (https://github.com/sahirbhatnagar/cbpaper.git)
#> Head:     [349d153] 2020-08-21: fix merge conflicts
\end{CodeOutput}
\end{CodeChunk}
